---
title: 'Reproducible Research: Peer Assessment 1'
author: "Gaurab Kundu"
date: "2022-10-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

It is required that data source file is already in a working directory. I check if the source file is unzipped and if it is not - I unzip it and load the data. Then I delete NA observations, since I do not need it for calculations.

```{r}
if (!file.exists("activity.csv") ) {
        unzip("activity.zip")
}
raw_data <- read.csv("activity.csv", header = TRUE)
main_data <- na.omit(raw_data)
```

Then I explore the data with functions ‘str’, ‘head’, ‘tail’, ‘table’ (for each column) in order to get aquianted with it.

## What is mean total number of steps taken per day?

```{r}
steps_per_day <- aggregate(main_data$steps, by = list(Steps.Date = main_data$date), FUN = "sum")
```

Then I make a histogram with the frequency ot total numbers

```{r}
hist(steps_per_day$x, col = "blue", 
     breaks = 20,
     main = "Total number of steps taken each day",
     xlab = "Number of steps per day")
```

And I calculate the mean and median of the total number of steps

```{r}
mean_steps <- mean(steps_per_day[,2])
print (mean_steps)
```

```{r}
median_steps <- median(steps_per_day[,2])
print (median_steps)
```

## What is the average daily activity pattern?

At first, we can look at the plot of the number of steps taken averaged across all days, along all 5-min intervals

```{r}
avaraged_day <- aggregate(main_data$steps, 
                          by = list(Interval = main_data$interval), 
                          FUN = "mean")
plot(avaraged_day$Interval, avaraged_day$x, type = "l", 
     main = "Average daily activity pattern", 
     ylab = "Avarage number of steps taken", 
     xlab = "5-min intervals")
```

Then we define the interval with the maximum number of steps

```{r}
interval_row <- which.max(avaraged_day$x)
max_interval <- avaraged_day[interval_row,1]
print (max_interval)
```

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?

