---
title: 'Reproducible Research: Peer Assessment 1'
author: "Gaurab Kundu"
date: "2022-10-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

It is required that data source file is already in a working directory. I check if the source file is unzipped and if it is not - I unzip it and load the data. Then I delete NA observations, since I do not need it for calculations.

```{r}
if (!file.exists("activity.csv") ) {
        unzip("activity.zip")
}
raw_data <- read.csv("activity.csv", header = TRUE)
main_data <- na.omit(raw_data)
```

Then I explore the data with functions ‘str’, ‘head’, ‘tail’, ‘table’ (for each column) in order to get aquianted with it.

## What is mean total number of steps taken per day?

```{r}
steps_per_day <- aggregate(main_data$steps, by = list(Steps.Date = main_data$date), FUN = "sum")
```

Then I make a histogram with the frequency ot total numbers

```{r}
hist(steps_per_day$x, col = "blue", 
     breaks = 20,
     main = "Total number of steps taken each day",
     xlab = "Number of steps per day")
```

And I calculate the mean and median of the total number of steps

```{r}
mean_steps <- mean(steps_per_day[,2])
print (mean_steps)
```

```{r}
median_steps <- median(steps_per_day[,2])
print (median_steps)
```


