---
title: 'Reproducible Research: Peer Assessment 1'
author: "Gaurab Kundu"
date: "2022-10-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

### It is required that data source file is already in a working directory. I check if the source file is unzipped and if it is not - I unzip it and load the data. Then I delete NA observations, since I do not need it for calculations.

```{r}
if (!file.exists("activity.csv") ) {
        unzip("activity.zip")
}
raw_data <- read.csv("activity.csv", header = TRUE)
main_data <- na.omit(raw_data)
```

### Then I explore the data with functions ‘str’, ‘head’, ‘tail’, ‘table’ (for each column) in order to get aquianted with it.

## What is mean total number of steps taken per day?

```{r}
steps_per_day <- aggregate(main_data$steps, by = list(Steps.Date = main_data$date), FUN = "sum")
```

### Then I make a histogram with the frequency ot total numbers

```{r}
hist(steps_per_day$x, col = "blue", 
     breaks = 20,
     main = "Total number of steps taken each day",
     xlab = "Number of steps per day")
```

### And I calculate the mean and median of the total number of steps

```{r}
mean_steps <- mean(steps_per_day[,2])
print (mean_steps)
```

```{r}
median_steps <- median(steps_per_day[,2])
print (median_steps)
```

## What is the average daily activity pattern?

### At first, we can look at the plot of the number of steps taken averaged across all days, along all 5-min intervals

```{r}
avaraged_day <- aggregate(main_data$steps, 
                          by = list(Interval = main_data$interval), 
                          FUN = "mean")
plot(avaraged_day$Interval, avaraged_day$x, type = "l", 
     main = "Average daily activity pattern", 
     ylab = "Avarage number of steps taken", 
     xlab = "5-min intervals")
```

### Then we define the interval with the maximum number of steps

```{r}
interval_row <- which.max(avaraged_day$x)
max_interval <- avaraged_day[interval_row,1]
print (max_interval)
```

## Imputing missing values

### I calculate the total number of NA values in the dataset

```{r}
NA_number <- length(which(is.na(raw_data$steps)))
print (NA_number)
```

### For substituting NA values I use simple ‘impute’ function in ‘Hmisc’ package (it is implied that this package is installed). I create a new dataset with the missing data filled in.

```{r}
library(Hmisc)
raw_data_filled <- raw_data
raw_data_filled$steps <- impute(raw_data$steps, fun=mean)
```

### Then I make histogram with the new frequencies of total number of steps after imputing.

```{r}
steps_per_day_noNA <- aggregate(raw_data_filled$steps, 
                                by = list(Steps.Date = raw_data_filled$date), 
                                FUN = "sum")
hist(steps_per_day_noNA$x, col = "green", 
     breaks = 20,
     main = "Total number of steps taken each day (filled data)",
     xlab = "Number of steps per day")
```

### And I calculate the new mean and median of the total number of steps after imputing.

```{r}
mean_steps_noNA <- mean(steps_per_day_noNA[,2])
print (mean_steps_noNA)
```

```{r}
median_steps_noNA <- median(steps_per_day_noNA[,2])
print (median_steps_noNA)
```

### As we can see, new figures are almost the same as before filling NA values. The median has changed a bit and became equal to the mean.

## Are there differences in activity patterns between weekdays and weekends?

